A Simple Minibatch Acceptance Test for MCMC

### 2016-05-17 nips_2016.tex
Replaced two figures with one nicer deconvolution visualization

\begin{figure}[t]
  \centering
  \includegraphics[width=\textwidth]{john_bair_fig01}
  \caption{
  To the left is the logistic distribution, representing a possible acceptance test. As $\Delta U$
  approaches infinity, the acceptance rate approaches 1. To the right, we have a new variable $X$
  and its density, along with an acceptance test $\Delta U + X$, which we describe in detail in
  Section~\ref{sec:our_algorithm} (note that we use $\Delta$ instead of $\Delta U$ for simplicity).
  {\color{blue}
  Daniel: I think that the diagram to the left should use the logistic function, i.e., replace
  ``$\exp(\Delta U)$'' with ``$(1+\exp(\Delta U))^{-1}$.''
  }
  }
  \label{fig:part1}
\end{figure}

\begin{figure}[t]
  \centering
  \includegraphics[width=\textwidth]{john_bair_fig02}
  \caption{
  To the left, we have $\Delta U$ (which is simplified as $\Delta$ in
  Section~\ref{sec:our_algorithm}), along with two noise terms that come from the new variable we
  define $X$. The details are in Section~\ref{sec:our_algorithm}.
  {\color{blue}
  Daniel: I think this needs to be merged into Figure~\ref{fig:part1}.
  }
  }
  \label{fig:part2}
\end{figure}

@@

For our MH test, the key is that we use a different $g$, the logistic function: $g(\Delta) =
(1+\exp(-\beta \Delta))^{-1}$. Again, straightforward arithmetic shows that it satisfies the
condition in Lemma~\ref{lem:detailed_balance}.  The logistic function is nice because we can easily
sample from it using the following procedure. At any iteration with a current parameter $\theta_t$
and a candidate sample $\theta'$, we can compute $\Delta$. Let $u$ be a uniform random variable $u
\sim {\rm Unif}[0,1]$. We accept $\theta'$ if $g(\Delta) > u$, and reject otherwise. This process is
equivalent to sampling a random variable $X$ with cumulative distribution function $F_X(x) = g(x)$,
and accepting if $\Delta > X$ and rejecting otherwise. To see why, define $X = g^{-1}(u)$ where
again, $g$ is the logistic function. The CDF of $X$ is precisely the logistic function because for
arbitrary $X = x$, we have
\[
F_X(x) = {\rm Pr}(X \le x) = {\rm Pr}(g^{-1}(u) \le x) = {\rm Pr}(u \le g(x)) = \int_{0}^{g(x)} 1 dx = g(x),
\]
as the density of the uniform random variable here is simply one. Thus, the criteria to accept the
candidate $\theta'$ is equivalent to whether $\Delta > X$. Since $X$ is symmetric about zero, the
acceptance criteria is equivalently expressed as $\Delta + X>0$, as shown in Figure~\ref{fig:part1}.

@@

From Lemma~\ref{lem:gaussian}, since $\Delta'$ is a noisy approximation of $\Delta$, the
relationship can be expressed as
\begin{equation}\label{eq:relationship}
\Delta' = \Delta + \varepsilon, \quad \varepsilon \sim \mathcal{N}(0, \sigma^2),
\end{equation}
for a Gaussian noise term $\varepsilon$.  The extra noise in $\Delta'$ means we can no longer perform
the $\Delta + X$ test, so we need to slightly change our acceptance criteria. We can decompose
$X$ as $X = X_{\rm norm}+X_{\rm corr}$, where the first term is a zero-mean ``noise term'' which
has the same variance as $\varepsilon$, and the second term is a zero-mean ``correction'' term.
These two convolve to form $X$. The criteria to accept can be rewritten as
\begin{equation}\label{eq:criteria}
\Delta + X = \Delta + X_{\rm norm} + X_{\rm corr} \approx \Delta' + X_{\rm corr} >0.
\end{equation}
This deconvolution process is visualized in Figure~\ref{fig:part2}. We decompose $X$ this way so
that we can estimate $X_{\rm norm}$ from the data, and then add the correction $X_{\rm corr}$ later.
Notice that we have some control over the true value of $X_{\rm norm}$, even though we cannot
generally guarantee computing it exactly. In particular, increasing the minibatch size will cause
the $X_{\rm norm}$ distribution to become narrower. {\color{blue} (Daniel: note that we will be
fixing this figure, so the visualization should hopefully be a little clearer, along with the
corresponding paragraph description. In particular, this paragraph needs an informal description of
Algorithm~\ref{alg:our_algorithm}.)} Algorithm~\ref{alg:our_algorithm} concisely describes our MH
test.

>>>>

\begin{figure}[t]
  \centering
  \includegraphics[width=\textwidth]{mh_convolution_diagram}
  \caption{
  {\color{blue}
  Daniel: this image is finalized, but change this caption.
  }
  }
  \label{fig:deconvolution}
\end{figure}

@@

For our MH test, the key is that we use a different $g$, the logistic function: $g(\Delta) =
(1+\exp(-\beta \Delta))^{-1}$. Again, straightforward arithmetic shows that it satisfies the
condition in Lemma~\ref{lem:detailed_balance}.  The logistic function is nice because we can easily
sample from it using the following procedure. At any iteration with a current parameter $\theta_t$
and a candidate sample $\theta'$, we can compute $\Delta$. Let $u$ be a uniform random variable $u
\sim {\rm Unif}[0,1]$. We accept $\theta'$ if $g(\Delta) > u$, and reject otherwise. This process is
equivalent to sampling a random variable $X$ with cumulative distribution function $F_X(x) = g(x)$,
and accepting if $\Delta > X$ and rejecting otherwise. To see why, define $X = g^{-1}(u)$ where
again, $g$ is the logistic function. The CDF of $X$ is precisely the logistic function because for
arbitrary $X = x$, we have
\[
F_X(x) = {\rm Pr}(X \le x) = {\rm Pr}(g^{-1}(u) \le x) = {\rm Pr}(u \le g(x)) = \int_{0}^{g(x)} 1 dx = g(x),
\]
as the density of the uniform random variable here is simply one. Thus, the criteria to accept the
candidate $\theta'$ is equivalent to whether $\Delta > X$. Since $X$ is symmetric about zero, the
acceptance criteria is equivalently expressed as $\Delta + X>0$.

@@

From Lemma~\ref{lem:gaussian}, since $\Delta'$ is a noisy approximation of $\Delta$, the
relationship can be expressed as
\begin{equation}\label{eq:relationship}
\Delta' = \Delta + \varepsilon, \quad \varepsilon \sim \mathcal{N}(0, \sigma^2),
\end{equation}
for a Gaussian noise term $\varepsilon$.  The extra noise in $\Delta'$ means we can no longer perform
the $\Delta + X$ test, so we need to slightly change our acceptance criteria. We can decompose
$X$ as $X = X_{\rm norm}+X_{\rm corr}$, where the first term is a zero-mean ``noise term'' which
has the same variance as $\varepsilon$, and the second term is a zero-mean ``correction'' term.
These two convolve to form $X$. The criteria to accept can be rewritten as
\begin{equation}\label{eq:criteria}
\Delta + X = \Delta + X_{\rm norm} + X_{\rm corr} \approx \Delta' + X_{\rm corr} >0.
\end{equation}
This deconvolution process is visualized in Figure~\ref{fig:deconvolution}. We decompose $X$ this
way so that we can estimate $X_{\rm norm}$ from the data, and then add the correction $X_{\rm corr}$
later.  Notice that we have some control over the true value of $X_{\rm norm}$, even though we
cannot generally guarantee computing it exactly. In particular, increasing the minibatch size will
cause the $X_{\rm norm}$ distribution to become narrower. {\color{blue} (Daniel: note that we will
be fixing this figure, so the visualization should hopefully be a little clearer, along with the
corresponding paragraph description. In particular, this paragraph needs an informal description of
Algorithm~\ref{alg:our_algorithm}.)} Algorithm~\ref{alg:our_algorithm} concisely describes our MH
test.
